---
title: "scRNAseq analysis of E-MTAB-8698_EMBOJ"
output: html_notebook
---
refer to:
Cattenoz PB, Sakr R, Pavlidaki A, Delaporte C, Riba A, Molina N, Hariharan N, Mukherjee T, Giangrande A. Temporal specificity and heterogeneity of Drosophila immune cells. EMBO J. 2020 Jun 17;39(12):e104486. doi: 10.15252/embj.2020104486. Epub 2020 Mar 12. PMID: 32162708; PMCID: PMC7298292.

```{r}

setwd(getwd())

library(Seurat)
library(ggplot2)
library(cowplot)
library(Matrix)
library(dplyr)
library(ggsci)
library(data.table)
library(monocle)
library(DoubletFinder)
library(tidyverse)
library(patchwork)
library(clustree)
library(remotes)




```


```{r}

meta_info <- data.table::fread("./E-MTAB-8698/E-MTAB-8698.sdrf.txt")

NI <- Seurat::Read10X("/Users/tangweihao/Desktop/summer2024_research/II_cancer2024summer/scRNAseq/E-MTAB-8698_EMBOJ/E-MTAB-8698/filtered_feature_bc_matrix_NI_RRCZ22")

NI <- CreateSeuratObject(counts = NI,
                           project = "Non_Infected_RRCZ22", 
                           min.cells = 0,  
                           min.features = 0)
```

```{r}
# 
# n.Feature.min <- 2  
# n.Feature.max <- 20000  
# n.Count.min <-2 
# n.Mt <- 5  
# n.Rb <- 5 

cat("Before filter :",nrow(NI@meta.data),"cells\n")



```

```{r}

# run standard anlaysis workflow
NI2 <- NormalizeData(NI)
NI2 <- FindVariableFeatures(NI2)
NI2 <- ScaleData(NI2)

min(nrow(NI2), ncol(NI2))
NI2 <- RunPCA(NI2,npcs = 50,verbose = F)

#see turning point
ElbowPlot(NI2, ndims=30, reduction="pca")


NI2 <- FindNeighbors(NI2, dims = 1:20, reduction = "pca", verbose=TRUE)

NI2 <- FindClusters(NI2, resolution = 1, cluster.name = "NI_clusters")

#see resolution
pdf(paste0("./cluster_tree.pdf"),width = 15,height = 10)
clustree(FindClusters(NI2, resolution = seq(0.6,1.2,by=0.05)))
dev.off()

NI2 <- FindClusters(NI2, resolution = 0.8, cluster.name = "NI_clusters")

```


```{r}
#Umap
NI3 <- RunUMAP(NI2, dims = 1:20, reduction = "pca", reduction.name = "UMAP_NI",verbose=F)

colnames(NI3@reductions[["UMAP_NI"]]@cell.embeddings) <- c('UMAP_1', 'UMAP_2')


DimPlot(NI3, reduction = "UMAP_NI", group.by = "orig.ident")

DimPlot(NI3, reduction = "UMAP_NI")




#tsne
NI3 <- RunTSNE(NI3, dims = 1:20, reduction = "pca", reduction.name = "tSNE_NI",verbose=F)

colnames(NI3@reductions[["tSNE_NI"]]@cell.embeddings) <- c('tSNE_1', 'tSNE_2')


DimPlot(NI3, reduction = "tSNE_NI", group.by = "orig.ident")

DimPlot(NI3, reduction = "tSNE_NI")
```

#find marker genes
#For clarification, refer to https://www.biostars.org/p/409790/
```{r}


test <- NI3

all.markers <- FindAllMarkers(test, test.use = 'wilcox', only.pos = TRUE, min.pct = 0.10, logfc.threshold = 0.25)

```

```{r}

marker.sig <- all.markers %>%
    filter(p_val_adj < 0.05)

topmarker <- marker.sig %>% group_by(cluster) %>% 
    top_n(n = 5, wt = avg_log2FC)

DoHeatmap(test, features = topmarker$gene,size = 2) +
    theme(legend.position = "none", 
          axis.text.y = element_text(size = 6))

DotPlot(test, features = unique(topmarker$gene))


# +
#     RotatedAxis()

pdf(paste0("./NI_DotPlot.pdf"),width = 15,height = 10)
DotPlot(test, features = unique(topmarker$gene))+
     theme(axis.text.x = element_text(angle = 90, size = 12), legend.text = element_text(size = 12), legend.title = element_text(size = 12))
dev.off()



pdf(paste0("./NI_Heatmap.pdf"),width = 15,height = 10)
DoHeatmap(test, features = topmarker$gene,size = 2) +
    theme(legend.position = "none", 
          axis.text.y = element_text(size = 6))
dev.off()


all_markers_dotplot <- DotPlot(test, features = unique(topmarker$gene))+
     theme(axis.text.x = element_text(angle = 90, size = 12), legend.text = element_text(size = 12), legend.title = element_text(size = 12))

View(all_markers_dotplot$data)

```


#feature plot of specific genes (for e.g. SPH93 in ScribKD and Ctrl)
```{r}
#split.by = "groups",
# FeaturePlot(test, features = c("SPH93","CecA1"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")
# 
# FeaturePlot(test, features = c("Rel","AttA"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")
#  



pdf(paste0("./NI_UMAP_DimPlot.pdf"),width = 15,height = 10)
DimPlot(test,reduction = "UMAP_NI",)
dev.off()



pdf(paste0("./NI_tSNE_DimPlot.pdf"),width = 15,height = 10)
DimPlot(test,reduction = "tSNE_NI")
dev.off()
```


#manual cell-cluster annotation based on Figure 2 D of https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7298292/figure/embj2020104486-fig-0002/

bsed on appendix figure S5
```{r}

#13# tau NLaz #PL-lmpl2#

#10# CecA1, CecA2, CecC #PL-AMP#

#checked#11# PPO1 PPO2 #CC#

#7# Hml Nplp2 Ppn #PL-prolif#

#2#  robo2 Ten-a #PL-robo2#

#8# #Inos #PL-Inos#

#12# E(spl)mbeta???HLH #PL-3#

#9# Lsp1alpha Lsp1beta Obp99b#PL_Lsp#

##  ##

```


#make a dotplot that includes just marker genes that are known to be specific to the crystal cells, to
the lamellocytes or to plasmatocytes , for each cluster
based on Appendix Figure S4
```{r}


pdf(paste0("./NI_DotPlot_Author_specific_markers.pdf"),width = 15,height = 10)
DotPlot(test, features = c("lz","peb","PPO2","PPO1","sn","Pxn","Hml","eater","NimC1","crq","He","srp"))+
     theme(axis.text.x = element_text(angle = 90, size = 16), legend.text = element_text(size = 12), legend.title = element_text(size = 12))
dev.off()


known_markers_dotplot <- DotPlot(test, features = c("lz","peb","PPO2","PPO1","sn","Pxn","Hml","eater","NimC1","crq","He","srp"))+
     theme(axis.text.x = element_text(angle = 90, size = 16), legend.text = element_text(size = 12), legend.title = element_text(size = 12))

View(known_markers_dotplot$data)

```

importing author's annotation
```{r}

Authors_markers <- openxlsx::read.xlsx("Authors_markers.xlsx",colNames = T,sheet = 2)

colnames(Authors_markers) <- Authors_markers[2,]

Authors_markers <- Authors_markers[-c(1:2),]




#cell types
Authors_celltypes <- unique(Authors_markers$cluster)

our_celltypes <- as.character(unique(all.markers$cluster))

Authors_celltypes
our_celltypes

# 
# our_celltypes_new <- our_celltypes
# 
# 
# for (i in 1:length(Authors_celltypes)) {
#   markers_common <- 
#     intersect(marker.sig$gene[which(marker.sig$cluster==our_celltypes[3])],
#         Authors_markers$"Official gene symbol"[which(Authors_markers$cluster==Authors_celltypes[i])]);
#   print(markers_common)
# }

# 
# our_celltypes_new[1] <- Authors_celltypes[14]
# our_celltypes_new[2] <- Authors_celltypes[7]
# our_celltypes_new[3] <- Authors_celltypes[14]
# our_celltypes_new[4] <- Authors_celltypes[]
# our_celltypes_new[5] <- Authors_celltypes[]
# our_celltypes_new[6] <- Authors_celltypes[]
# our_celltypes_new[7] <- Authors_celltypes[]
# our_celltypes_new[8] <- Authors_celltypes[]
# our_celltypes_new[9] <- Authors_celltypes[]
# our_celltypes_new[10] <- Authors_celltypes[]
# our_celltypes_new[11] <- Authors_celltypes[]
# our_celltypes_new[12] <- Authors_celltypes[1]
# our_celltypes_new[13] <- Authors_celltypes[]
# our_celltypes_new[14] <- Authors_celltypes[]
# 
# 
# our_celltypes_new





cluster2celltype <- c("0"="PL-0",
                      "1"="PL-1",
                      "2"="PL-robo2", 
                      "3"="PL-Rel", 
                      "4"="PL-Inos",
                      "5"="PL-3",
                      "6"="PL-Pcd",
                      "7"="PL-vir1", 
                      "8"="PL-prolif",
                      "9"="PL-Lsp",
                      "10"="CC",
                      "11"="PL-AMP",
                      "12"="PL-2",
                      "13"="PL-ImpL2")

test[['cell_type']] <- unname(cluster2celltype[test@meta.data$seurat_clusters])


#plot order of clusters (to match authors)
# test$cell_type<- factor(test$cell_type,levels = c(
#   "",
#   "",
#   "",
#   "",
#   "",
#   "",
#   "",
#   "",
#   "",
#   "",
#   "",
#   "",
#   "",
#   "",))

```



#checking if there is swap comparing to authors' data
```{r}

DimPlot(test,reduction = "UMAP_NI",group.by = "cell_type")
DimPlot(test,reduction = "tSNE_NI",group.by = "cell_type")


pdf(paste0("./DimPlot_UMAP_NI_clusterNamed.pdf"),width = 15,height = 10)
DimPlot(test,reduction = "UMAP_NI",group.by = "cell_type",
        label = TRUE,
        label.size = 3,
        repel = TRUE)
dev.off()


pdf(paste0("./DimPlot_tSNE_NI_clusterNamed.pdf"),width = 15,height = 10)
DimPlot(test,reduction = "tSNE_NI",group.by = "cell_type",
        label = TRUE,
        label.size = 3,
        repel = TRUE)
dev.off()


#Umap feature plot
FeaturePlot(test, features = c("Pcd"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("Inos"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("ImpL2"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("Hml"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")



FeaturePlot(test, features = c("vir-1"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("Rel"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("sn"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("robo2"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")


#tSNE feature plot

FeaturePlot(test, features = c("Pcd"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("Inos"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("ImpL2"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("Hml"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")



FeaturePlot(test, features = c("vir-1"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("Rel"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("sn"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("robo2"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

```





#feature plot of genes of lab interest
```{r}

#UMAP
DimPlot(test,reduction = "UMAP_NI",group.by = "cell_type")

FeaturePlot(test, features = c("SPH93","CecA1"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("Rel","AttA"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("AttB","edin"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")
FeaturePlot(test, features = c("CecA2","Abl"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("AttC","DptA"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("DptB","Dro"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")


FeaturePlot(test, features = c("Mtk","IM18"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")


#tSNE


DimPlot(test,reduction = "tSNE_NI",group.by = "cell_type")

FeaturePlot(test, features = c("SPH93","CecA1"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("Rel","AttA"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("AttB","edin"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")
FeaturePlot(test, features = c("CecA2","Abl"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("AttC","DptA"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("DptB","Dro"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")


FeaturePlot(test, features = c("Mtk","IM18"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")


```

plot dotplot, heatmap for named clusters
```{r}


pdf(paste0("./NI_DotPlot_Author_specific_markers_clusterNamed.pdf"),width = 15,height = 10)
DotPlot(test, group.by = "cell_type",features = c("lz","peb","PPO2","PPO1","sn","Pxn","Hml","eater","NimC1","crq","He","srp"))+
     theme(axis.text.x = element_text(angle = 90, size = 16), legend.text = element_text(size = 12), legend.title = element_text(size = 12))
dev.off()


pdf(paste0("./NI_DotPlot_clusterNamed.pdf"),width = 15,height = 10)
DotPlot(test, group.by = "cell_type",features = unique(topmarker$gene))+
     theme(axis.text.x = element_text(angle = 90, size = 12), legend.text = element_text(size = 12), legend.title = element_text(size = 12))
dev.off()



pdf(paste0("./NI_Heatmap_clusterNamed.pdf"),width = 15,height = 10)
DoHeatmap(test, features = topmarker$gene,size = 2,group.by = "cell_type") +
    theme(legend.position = "none", 
          axis.text.y = element_text(size = 6))
dev.off()

```
```{r}

library(grid)

DimPlot(test,reduction = "tSNE_NI",group.by = "cell_type")

rasterImage(
  FeaturePlot(test, features = c("CecA1"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")
  )


```

