---
title: "scRNAseq analysis of E-MTAB-8698_EMBOJ"
output: html_notebook
---
#refer to:
Cattenoz PB, Sakr R, Pavlidaki A, Delaporte C, Riba A, Molina N, Hariharan N, Mukherjee T, Giangrande A. Temporal specificity and heterogeneity of Drosophila immune cells. EMBO J. 2020 Jun 17;39(12):e104486. doi: 10.15252/embj.2020104486. Epub 2020 Mar 12. PMID: 32162708; PMCID: PMC7298292.

#objective:
## 1. Try to reproduce Dimplot of author's Fig. 2A in the paper
## 2. See in which kind of cells are III markers expressed in hemocytes

#Important files***:
## 1. E-MTAB-8698.sdrf.txt meta information for seurat object
## 2. filtered_feature_bc_matrix_NI_RRCZ22, unzipped folder that contains control (Non_Infected) group of hemocytes scRNAseq data
*** These files can be downloaded at https://www.ebi.ac.uk/biostudies/arrayexpress/studies/E%E2%80%90MTAB%E2%80%908698

#######################################
set wk dir and library package
```{r}
setwd(getwd())

library(Seurat)
library(ggplot2)
library(cowplot)
library(Matrix)
library(dplyr)
library(ggsci)
library(data.table)
library(tidyverse)
library(patchwork)
library(clustree)
library(remotes)

```

#import important files and create seurat object
```{r}

meta_info <- data.table::fread("./E-MTAB-8698/E-MTAB-8698.sdrf.txt")

NI <- Seurat::Read10X("/Users/tangweihao/Desktop/summer2024_research/II_cancer2024summer/scRNAseq/E-MTAB-8698_EMBOJ/E-MTAB-8698/filtered_feature_bc_matrix_NI_RRCZ22")

NI <- CreateSeuratObject(counts = NI,
                           project = "Non_Infected_RRCZ22", 
                           min.cells = 0,  
                           min.features = 0)
```

#quality control
```{r}
# 
# n.Feature.min <- 2  
# n.Feature.max <- 20000  
# n.Count.min <-2 
# n.Mt <- 5  
# n.Rb <- 5 

cat("Before filter :",nrow(NI@meta.data),"cells\n")



```

#run standard seurat pipeline
```{r}

# run standard anlaysis workflow
NI2 <- NormalizeData(NI)
NI2 <- FindVariableFeatures(NI2)
NI2 <- ScaleData(NI2)

min(nrow(NI2), ncol(NI2))
NI2 <- RunPCA(NI2,npcs = 50,verbose = F)

#see turning point
ElbowPlot(NI2, ndims=30, reduction="pca")


NI2 <- FindNeighbors(NI2, dims = 1:20, reduction = "pca", verbose=TRUE)

NI2 <- FindClusters(NI2, resolution = 1, cluster.name = "NI_clusters")

#see resolution
pdf(paste0("./cluster_tree.pdf"),width = 15,height = 10)
clustree(FindClusters(NI2, resolution = seq(0.6,1.2,by=0.05)))
dev.off()

NI2 <- FindClusters(NI2, resolution = 0.8, cluster.name = "NI_clusters")

```

#Run UMAP and tSNE dim reduction
```{r}
#Umap
NI3 <- RunUMAP(NI2, dims = 1:20, reduction = "pca", reduction.name = "UMAP_NI",verbose=F)

colnames(NI3@reductions[["UMAP_NI"]]@cell.embeddings) <- c('UMAP_1', 'UMAP_2')


DimPlot(NI3, reduction = "UMAP_NI", group.by = "orig.ident")

DimPlot(NI3, reduction = "UMAP_NI")




#tsne
NI3 <- RunTSNE(NI3, dims = 1:20, reduction = "pca", reduction.name = "tSNE_NI",verbose=F)

colnames(NI3@reductions[["tSNE_NI"]]@cell.embeddings) <- c('tSNE_1', 'tSNE_2')


DimPlot(NI3, reduction = "tSNE_NI", group.by = "orig.ident")

DimPlot(NI3, reduction = "tSNE_NI")
```

#find marker genes
#For clarification, refer to https://www.biostars.org/p/409790/
```{r}

test <- NI3

all.markers <- FindAllMarkers(test, test.use = 'wilcox', only.pos = TRUE, min.pct = 0.10, logfc.threshold = 0.25)

```

#generate dotplot to compare with author's Fig 2.D
```{r}

marker.sig <- all.markers %>%
    filter(p_val_adj < 0.05)

topmarker <- marker.sig %>% group_by(cluster) %>% 
    top_n(n = 5, wt = avg_log2FC)

DoHeatmap(test, features = topmarker$gene,size = 2) +
    theme(legend.position = "none", 
          axis.text.y = element_text(size = 6))

DotPlot(test, features = unique(topmarker$gene))


# +
#     RotatedAxis()

pdf(paste0("./NI_DotPlot.pdf"),width = 15,height = 10)
DotPlot(test, features = unique(topmarker$gene))+
     theme(axis.text.x = element_text(angle = 90, size = 12), legend.text = element_text(size = 12), legend.title = element_text(size = 12))
dev.off()



pdf(paste0("./NI_Heatmap.pdf"),width = 15,height = 10)
DoHeatmap(test, features = topmarker$gene,size = 2) +
    theme(legend.position = "none", 
          axis.text.y = element_text(size = 6))
dev.off()


all_markers_dotplot <- DotPlot(test, features = unique(topmarker$gene))+
     theme(axis.text.x = element_text(angle = 90, size = 12), legend.text = element_text(size = 12), legend.title = element_text(size = 12))

View(all_markers_dotplot$data)

```


#feature plot of specific genes (for e.g. SPH93 in Ctrl)
```{r}
#split.by = "groups",
# FeaturePlot(test, features = c("SPH93","CecA1"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")
# 
# FeaturePlot(test, features = c("Rel","AttA"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")
#  



pdf(paste0("./NI_UMAP_DimPlot.pdf"),width = 15,height = 10)
DimPlot(test,reduction = "UMAP_NI",)
dev.off()



pdf(paste0("./NI_tSNE_DimPlot.pdf"),width = 15,height = 10)
DimPlot(test,reduction = "tSNE_NI")
dev.off()
```


#manual cell-cluster annotation based on Figure 2 D of https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7298292/figure/embj2020104486-fig-0002/

bsed on appendix figure S5
```{r}

#13# tau NLaz #PL-lmpl2#

#10# CecA1, CecA2, CecC #PL-AMP#

#checked#11# PPO1 PPO2 #CC#

#7# Hml Nplp2 Ppn #PL-prolif#

#2#  robo2 Ten-a #PL-robo2#

#8# #Inos #PL-Inos#

#12# E(spl)mbeta???HLH #PL-3#

#9# Lsp1alpha Lsp1beta Obp99b#PL_Lsp#

##  ##

```


#make a dotplot that includes just marker genes that are known to be specific to the crystal cells, to
the lamellocytes or to plasmatocytes , for each cluster
based on Appendix Figure S4
```{r}


pdf(paste0("./NI_DotPlot_Author_specific_markers.pdf"),width = 15,height = 10)
DotPlot(test, features = c("lz","peb","PPO2","PPO1","sn","Pxn","Hml","eater","NimC1","crq","He","srp"))+
     theme(axis.text.x = element_text(angle = 90, size = 16), legend.text = element_text(size = 12), legend.title = element_text(size = 12))
dev.off()


known_markers_dotplot <- DotPlot(test, features = c("lz","peb","PPO2","PPO1","sn","Pxn","Hml","eater","NimC1","crq","He","srp"))+
     theme(axis.text.x = element_text(angle = 90, size = 16), legend.text = element_text(size = 12), legend.title = element_text(size = 12))

View(known_markers_dotplot$data)

```

#change cluster ID to cell type
```{r}

cluster2celltype <- c("0"="PL-0",
                      "1"="PL-1",
                      "2"="PL-robo2", 
                      "3"="PL-Rel", 
                      "4"="PL-Inos",
                      "5"="PL-3",
                      "6"="PL-Pcd",
                      "7"="PL-vir1", 
                      "8"="PL-prolif",
                      "9"="PL-Lsp",
                      "10"="CC",
                      "11"="PL-AMP",
                      "12"="PL-2",
                      "13"="PL-ImpL2")

test[['cell_type']] <- unname(cluster2celltype[test@meta.data$seurat_clusters])

```



#checking if there is swap comparing to authors' data
```{r}

DimPlot(test,reduction = "UMAP_NI",group.by = "cell_type")
DimPlot(test,reduction = "tSNE_NI",group.by = "cell_type")


pdf(paste0("./DimPlot_UMAP_NI_clusterNamed.pdf"),width = 15,height = 10)
DimPlot(test,reduction = "UMAP_NI",group.by = "cell_type",
        label = TRUE,
        label.size = 3,
        repel = TRUE)
dev.off()


pdf(paste0("./DimPlot_tSNE_NI_clusterNamed.pdf"),width = 15,height = 10)
DimPlot(test,reduction = "tSNE_NI",group.by = "cell_type",
        label = TRUE,
        label.size = 3,
        repel = TRUE)
dev.off()


#Umap feature plot
FeaturePlot(test, features = c("Pcd"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("Inos"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("ImpL2"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("Hml"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")



FeaturePlot(test, features = c("vir-1"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("Rel"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("sn"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("robo2"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")


#tSNE feature plot

FeaturePlot(test, features = c("Pcd"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("Inos"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("ImpL2"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("Hml"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")



FeaturePlot(test, features = c("vir-1"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("Rel"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("sn"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("robo2"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

```





#feature plot of genes of interest
```{r}

#UMAP
DimPlot(test,reduction = "UMAP_NI",group.by = "cell_type")

FeaturePlot(test, features = c("SPH93","CecA1"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("Rel","AttA"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("AttB","edin"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")
FeaturePlot(test, features = c("CecA2","Abl"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("AttC","DptA"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")

FeaturePlot(test, features = c("DptB","Dro"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")


FeaturePlot(test, features = c("Mtk","IM18"),  cols = c("lightblue2","darkred"), reduction = "UMAP_NI",keep.scale = "feature")


#tSNE


DimPlot(test,reduction = "tSNE_NI",group.by = "cell_type")

FeaturePlot(test, features = c("SPH93","CecA1"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("Rel","AttA"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("AttB","edin"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")
FeaturePlot(test, features = c("CecA2","Abl"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("AttC","DptA"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")

FeaturePlot(test, features = c("DptB","Dro"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")


FeaturePlot(test, features = c("Mtk","IM18"),  cols = c("lightblue2","darkred"), reduction = "tSNE_NI",keep.scale = "feature")


```

#plot dotplot, heatmap for named clusters
```{r}


pdf(paste0("./NI_DotPlot_Author_specific_markers_clusterNamed.pdf"),width = 15,height = 10)
DotPlot(test, group.by = "cell_type",features = c("lz","peb","PPO2","PPO1","sn","Pxn","Hml","eater","NimC1","crq","He","srp"))+
     theme(axis.text.x = element_text(angle = 90, size = 16), legend.text = element_text(size = 12), legend.title = element_text(size = 12))
dev.off()


pdf(paste0("./NI_DotPlot_clusterNamed.pdf"),width = 15,height = 10)
DotPlot(test, group.by = "cell_type",features = unique(topmarker$gene))+
     theme(axis.text.x = element_text(angle = 90, size = 12), legend.text = element_text(size = 12), legend.title = element_text(size = 12))
dev.off()



pdf(paste0("./NI_Heatmap_clusterNamed.pdf"),width = 15,height = 10)
DoHeatmap(test, features = topmarker$gene,size = 2,group.by = "cell_type") +
    theme(legend.position = "none", 
          axis.text.y = element_text(size = 6))
dev.off()

```

